---
title: "Data acquisition and preparation"
author: "Jan Caha"
output: html_document
---

## Set up R and necessary packages

```{r setup, message = FALSE, warning = FALSE}
source("../scripts/_init.R")
walk(list.files(here("functions"), full.names = TRUE), source)
```

## Download data


### Obtain Package list

The data were downloaded on 9. 8. 2019. The code is provided here, but all the subsequent analyses are done on the saved data.

```{r}
# packages <- httr::content(httr::GET("http://crandb.r-pkg.org/-/desc"))
# 
# packages_info <- tibble(package = names(packages))
# 
# packages_info <- packages_info %>%
#   mutate(version = map_chr(packages, function(x) x[[1]]),
#          title = map_chr(packages, function(x) x[[2]]))
# 
# rm(packages)
# 
# packages_info %>% 
#   write_rds(here::here("data", "produced", "packages_basic_info.Rdata"))
```

### Get Metadata for all packages 

In ideal world this code snippet would do the work, but since it is significant number of API queries it often fails as a result of network timeout.

```{r}
# packages_info <- packages_info %>%
#   mutate(metadata = map(package, function(x) {package(x)}))
```

This code is less elegant but more practical. If the `for` loop fails, it can be rerun from the last succesful part, if the value of variable `i` was not reseted. This leads way the data scrapping is more often succesul. The `Sys.sleep()` command is used to ensure that we do not overload the target server. Since this is time consuming step a textual progress bar is provided.

```{r}
# list_metadata <- list()
# block_size <- 500
# i <- 1
# 
# pb <- txtProgressBar(min = 1, max = nrow(packages_info))
# 
# while (i < nrow(packages_info)) {
# 
#   setTxtProgressBar(pb, i)
# 
#   data <- packages_info %>% slice(i:(i + block_size - 1))
#   list_to_append <- map(data$package, function(x) {package(x)})
#   list_metadata <- append(list_metadata, list_to_append)
#   i <- i + block_size
# 
#   Sys.sleep(2)
# }
# 
# close(pb)
# 
# list_metadata %>% 
#   write_rds(here::here("data", "produced", "metadata_all_packages.Rdata"))
```

## Determine package presence in CRAN Task Views

Main step is to scrap the list CRAN views and then from each view a list of packages.

```{r}
# link = "https://cran.r-project.org/web/views/"
# CRAN_views_list <- read_html(link)
# 
# views <- CRAN_views_list %>%
#   html_nodes("table") %>%
#   .[[1]] %>%
#   html_nodes("tr") %>%
#   html_node("a") %>%
#   html_text()
# 
# views_links <- CRAN_views_list %>%
#   html_nodes("table") %>%
#   .[[1]] %>%
#   html_nodes("tr") %>%
#   html_node("a") %>%
#   html_attr("href")
# 
# CRAN_views <- bind_cols(view = views, view_local_link = views_links) %>%
#   mutate(link = paste0(link, view_local_link))
# 
# rm(CRAN_views_list, views, views_links)
# 
# package_views <- tibble(package = character(), view = character())
# 
# for (i in seq_len(nrow(CRAN_views))) {
# 
#   packages <- read_html(CRAN_views$link[i]) %>%
#     html_nodes("ul") %>%
#     .[[length(.) - 1]] %>%
#     html_nodes("li") %>%
#     html_node("a") %>%
#     html_text()
# 
#   package_views <- bind_rows(package_views, tibble(package = packages, view = CRAN_views$view[i]))
# }
# 
# rm(i, link)
# 
# package_views %>% 
#   write_rds(here::here("data", "produced", "package_cran_views.RData"))
```

## Load the preprepared data

```{r}
packages_info <- read_rds(here::here("data", "produced", "packages_basic_info.Rdata"))
  
list_metadata <- read_rds(here::here("data", "produced", "metadata_all_packages.Rdata"))

package_views <- read_rds(here::here("data", "produced", "package_cran_views.RData"))
```


## Merge metadata with packages

```{r}
packages_info  <- packages_info %>% 
  mutate(metadata = list_metadata)

rm(list_metadata)
```

## Extract data from Metadata list column to specific columns

```{r}
packages_info <- packages_info %>% 
  mutate(description = map_chr(metadata, function(x) {str_replace_all(x$Description, "\n", " ")} ),
         imports_depends = map(metadata, function(x) {append(as.list(names(x$Imports)),
                                                             as.list(names(x$Depends)))}))
```

## Remove unnecessary columns 

```{r}
packages_info <- packages_info %>% 
  select(package, description, imports_depends)
```

## Process CRAN Views information

For our topic the views [Spatial](https://cran.r-project.org/web/views/Spatial.html) and [SpatioTemporal](https://cran.r-project.org/web/views/SpatioTemporal.html) are relevant. If package is listed under either of these CRAN Views then the package is always assigned to a special view "SpatialViews".

```{r}
package_views <- package_views %>% 
  mutate(view = case_when(
    view ==  "Spatial" ~ "SpatialViews",
    view == "SpatioTemporal" ~ "SpatialViews",
    TRUE ~ view)
  )
```


```{r}
spatial_packages_names <- package_views %>% 
  filter(view == "SpatialViews") %>% 
  pull(package) %>% 
  unique()
```

### Merging information about packages and CRAN Views

The presence of package in CRAN Views can be linked to other package information. The packages not listed in any CRAN View are marked as "NotListed".

```{r}
packages_info <- packages_info %>% 
  left_join(package_views, by = "package")

packages_info <- packages_info %>% 
  mutate(view = if_else(is.na(view), "NotListed", view))
```

Number of unique Views listed is `r packages_info$view %>% unique() %>% sort() %>% length()`.

## Storing the output data

Replace unicode symbols (unwanted and irrelevant) in package description with empty space.

```{r}
packages_info <- packages_info %>% 
  mutate(description = str_replace_all(description, "<[U|\\+|0-9|a-z]+>", " "))
```

The resulting datasets are saved for further use.

Unique list of packages, CRAN View is the first one that occurs unless it is "SpatialViews", that is listed allways under this specific view.

```{r}
packages_info %>% 
  distinct(package, .keep_all = TRUE) %>% 
  mutate(view = if_else(package %in% spatial_packages_names, "SpatialViews", view)) %>% 
  write_rds(here::here("data", "produced", "packages_unique.Rdata"))
```

Packages and respective views, some packages listed more than once.

```{r}
packages_info %>% 
  write_rds(here::here("data", "produced", "packages_info_for_analysis.Rdata"))
```
