---
title: "Evaluation of packages that depend on spatial package or contain domain specific word"
author: "Jan Caha"
output: html_document
---

```{r setup, message = FALSE, warning = FALSE}
source("../scripts/_init.R")

walk(list.files(here("functions"), full.names = TRUE), source)

packages_all <- read_rds(here("data", "produced", "packages_evaluation.Rdata"))

packages_not_spatial <- packages_all %>%
  filter(view != "SpatialViews")

column_names_found_packages <- c("Package", "Contains Domain Words", "Depends on Spatial Packages")
column_names_spatial_packages <- c("Package")

# cran_link <- function(x) {
#   paste0('<a href ="https://cran.r-project.org/web/packages/', 
#          x, 
#          '/index.html">',
#          x, 
#          '</a>')
# }
# 
# packages_all <- packages_not_spatial %>% 
#   mutate(link = map_chr(package, cran_link))
```

There are `r nrow(packages_all)` and `r nrow(packages_not_spatial)` outside of SpatialViews.

## Packages found based on dependence or text analysis

In this section we look at all the packages not listed under __Spatial__ or __SpatioTemporal__ CRAN View that can be found based on dependency on important spatial package or by containing domain specific word.

### Packages that contain domain words in description

```{r}
packages_not_spatial %>%
  filter(contains_domain_words == TRUE) %>% 
  select(package, contains_domain_words, spatial_package_dependency) %>% 
  DT::datatable(escape = FALSE,
                colnames = column_names_found_packages,
                rownames = FALSE)
```

### Packages that depend on the important spatial packages

```{r}
packages_not_spatial %>%
  filter(spatial_package_dependency == TRUE) %>%  
  select(package, contains_domain_words, spatial_package_dependency) %>%
  DT::datatable(escape = FALSE,
                colnames = column_names_found_packages,
                rownames = FALSE)
```

### Packages that depend on important spatial packages and contain the domain words

```{r}
packages_not_spatial %>% 
  filter(spatial_package_dependency == TRUE, 
         contains_domain_words == TRUE) %>% 
  select(package, contains_domain_words, spatial_package_dependency) %>%
  DT::datatable(escape = FALSE,
                colnames = column_names_found_packages,
                rownames = FALSE)
```

## Counts


There are `r packages_not_spatial %>% filter(spatial_package_dependency == TRUE, contains_domain_words == TRUE) %>% count() %>% pull(n)` packages that were found based on both identification criteria - dependence on important spatial package and occurrence of domain word in description.

There are `r packages_not_spatial %>% filter(spatial_package_dependency == TRUE) %>% count() %>% pull(n)` that have dependency on at least one important spatial package.

`r packages_not_spatial %>% filter(contains_domain_words == TRUE) %>% count() %>% pull(n)` contain domain specific word from geoinformatics in description.


At least one of the criteria meets `r packages_not_spatial %>% filter(contains_domain_words == TRUE | spatial_package_dependency == TRUE) %>% count() %>% pull(n)` packages. `r packages_not_spatial %>% filter(spatial_package_dependency == TRUE | contains_domain_words == TRUE,view != "NotListed") %>% count() %>% pull(n)` of these is in some other CRAN View besides Spatial and SpatioTemporal.

## Samples

Setting seed to make the selection of sample reproducible. 

```{r}
set.seed(485931)
```

**Examples for the manuscript.**

```{r}
sample <- packages_not_spatial %>% 
  filter(contains_domain_words == TRUE) %>% 
  sample_n(3) %>% 
  select(package, description)

sample %>% 
  DT::datatable(escape = FALSE,
                colnames = c("Package", "Description"),
                rownames = FALSE)
```

```{r}
sample <- packages_not_spatial %>% 
  filter(spatial_package_dependency == TRUE) %>% 
  sample_n(3) %>% 
  mutate(imports_depends = purrr::map_chr(imports_depends, glue::glue_collapse, sep = ", ")) %>% 
  select(package, imports_depends)

sample %>% 
  DT::datatable(escape = FALSE,
                colnames = c("Package", "Dependency"),
                rownames = FALSE)
```
